Towards caring 15-minute neighbourhoods

### 2024-05-16 paper/Manuscript.qmd
Tidying up the repo

---
title: "Towards caring 15-minute neighbourhoods"
author:
  - name: Anastasia Soukhov 
    email: soukhoa@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
    attributes:
        corresponding: true
  - name: Léa Ravensbergen
    email: ravensbl@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
  - name: Lucía Mejía Dorantes
    email: lucia.mejia.dorantes@gmail.com
    affiliations:
        - name: Consultant
          city: Karlsruhe, Germany
  - name: Antonio Páez
    email: paezha@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
format: 
  docx
  # elsevier-pdf:
  #   keep-tex: true
  #   include-in-header:
  #     text: |
  #       \usepackage{graphicx}
  #       \usepackage{unicode-math}
  #       \usepackage{hyperref}
  #       \def\tightlist{}
  #       \usepackage{setspace}
  #       \doublespacing
  #       \usepackage{lineno}
  #       \linenumbers
  #   journal:
  #     name: 7th International Conference on Women and Gender in Transportation (WGiT)
  #     formatting: preprint
  #     model: 3p
  #     cite-style: authoryear
keywords: ["Accessibility", "Mobility of Care", "Gender", "15-Minute City", "Care Complete Neighbourhoods"]
bibliography: bibliography.bib
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      cache=FALSE)
```

```{r}
#function for creating comments
word_comment <- function(comment, highlight = "") {
  if (isTRUE(knitr:::pandoc_to() == "docx")) {
    paste0('[', comment, ']{.comment-start id="0" author="Anastasia"',
           'date="2024-05-16T15:00:00Z"}', highlight, '[]{.comment-end id="0"}')
  }
}
```

```{r library-setup}
library(biscale)
library(cowplot)
library(sf)
library(here)
library(dplyr)
library(tmap)
library(mapview)
library(tmaptools)
library(pander)
library(flextable)
library(ftExtra)
library(scales)
```

```{r}
load("data/care_dest.rda")
```

@@

- Next: describe hamilton's land-use context; communities which are urban, rural, suburban. Also describe the residential parcels summary statistics, ~140,000 points, aggregated at the level of Canadian Census DA. See population descriptive plots in @fig-Fig1:
```{r plot-figure1, out.width=600, out.height=600}
#| label: fig-Fig1
#| fig-cap: "The number of residential parcels per DA in 2020 (top) and the rate of population per dwelling per DA (middle) and prevalence of Low-income cut-offs after-tax (LICO-AT) per DA (bottom) retrieved from the 2021 Canadian Census. All scales in quartiles. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics("data/pop_descriptive_plots.tiff")
```

@@

```{r plot-figure2, out.width=600, out.height=600}
#| label: fig-Fig2
#| fig-cap: "The locations of care destinations in Hamilton separated by the author-generated categories of: dependent-, errand-, grocery- and health- centric care categories. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics("data/care_categories_plot.tiff")
```

@@

```{r plot-figure3, out.width=600, out.height=600}
#| label: fig-Fig3
#| fig-cap: "The cumulative opportunity measure. The number of care destinations that can be reached, per DA, within 15 min walk. Visualised as 'caring' (based on the opportunity reachable number; Caring is approximately the median) and if complete (if all care categories can be reached; this is an arbitrary selection). Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics("data/all_cats_copp_weighted_plot.tiff")
```

```{r plot-figure4, out.width=600, out.height=600}
#| label: fig-Fig4
#| fig-cap: "The diversity measure based on the _specific type_ of destination cumulative opportunity DA measure. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics("data/all_cats_entropy_copp_weighted_plot.tiff")
```

```{r plot-figure5, out.width=600, out.height=600}
#| label: fig-Fig5
#| fig-cap: "A bivariate map of the cumulative opportuninities map (aggregated per category at DA - fig3) and the diversity measure (fig 4). Amenity diverse = diviersity index above the region median. Caring/Complete = accessibility value above the region median. The Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics("data/all_cats_bivariate_entropy_copp_plot.tiff")
```

BONUS! Per capita spatial availability. This is not discussed in the methods. The spatial availability calculation assumes population is 'population density' and weighted opportunity count. With this map we can start thinking about per capita availability of opportunities. 
```{r plot-figure6, out.width=600, out.height=600}
#| label: fig-Fig6
#| fig-cap: "Per capita spatial availability of care opportunities. Uses weighted care opportunities, population density as competition, and 15 minute binary cut-off. Aggregates parcel level spatial availability at the DA level (by summation) and divides by the DA population. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics("data/all_cats_vvi_plot.tiff")
```

@@

```{r plot-figure7, out.width=600, out.height=600}
#| label: fig-Fig7
#| fig-cap: "UNWEIGHTED -- The cumulative opportunity measure. The number of care destinations that can be reached, per DA, within 15 min walk. Visualised as 'caring' (based on the opportunity reachable number; Caring is approximately the median) and if complete (if all care categories can be reached; this is an arbitrary selection). Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics("data/all_cats_copp_plot.tiff")
```

```{r plot-figure8, out.width=600, out.height=600}
#| label: fig-Fig8
#| fig-cap: "UNWEIGHTED -- The diversity measure based on the _specific type_ of destination cumulative opportunity DA measure. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics("data/all_cats_entropy_copp_plot.tiff")
```

>>>>

---
title: "Towards caring 15-minute neighbourhoods"
author:
  - name: Anastasia Soukhov 
    email: soukhoa@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
    attributes:
        corresponding: true
  - name: Léa Ravensbergen
    email: ravensbl@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
  - name: Lucía Mejía Dorantes
    email: lucia.mejia.dorantes@gmail.com
    affiliations:
        - name: Consultant
          city: Karlsruhe, Germany
  - name: Antonio Páez
    email: paezha@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
format: 
  docx
  # elsevier-pdf:
  #   keep-tex: true
  #   include-in-header:
  #     text: |
  #       \usepackage{graphicx}
  #       \usepackage{unicode-math}
  #       \usepackage{hyperref}
  #       \def\tightlist{}
  #       \usepackage{setspace}
  #       \doublespacing
  #       \usepackage{lineno}
  #       \linenumbers
  #   journal:
  #     name: 7th International Conference on Women and Gender in Transportation (WGiT)
  #     formatting: preprint
  #     model: 3p
  #     cite-style: authoryear
keywords: ["Accessibility", "Mobility of Care", "Gender", "15-Minute City", "Care Complete Neighbourhoods"]
bibliography: bibliography.bib
---
```{r}
#| label: setup-knitr-options
#| include: false

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      cache=FALSE)

```
```{r }
#| label: set-up-copy-bibliography
#| include: false

# WARNING: DO NOT MODIFY THE BIBLIOGRAPHY ON THIS FOLDER, WORK WITH THE ONE IN FOLDER ./bibliography/
file.copy(glue::glue(here::here(), "/bibliography/bibliography.bib"), 
          getwd(),
          overwrite = TRUE)
```

```{r}
#| label: function for creating comments

word_comment <- function(comment, highlight = "") {
  if (isTRUE(knitr:::pandoc_to() == "docx")) {
    paste0('[', comment, ']{.comment-start id="0" author="Anastasia"',
           'date="2024-05-16T15:00:00Z"}', highlight, '[]{.comment-end id="0"}')
  }
}
```

```{r library-setup}
library(biscale)
library(cowplot)
library(dplyr)
library(flextable)
library(ftExtra)
library(glue)
library(here)
library(mapview)
library(pander)
library(scales)
library(sf)
library(tmap)
library(tmaptools)
```

```{r}
load(glue::glue(here::here(), "/data/care_dest.rda"))
```

@@

- Next: describe hamilton's land-use context; communities which are urban, rural, suburban. Also describe the residential parcels summary statistics, ~140,000 points, aggregated at the level of Canadian Census DA. See population descriptive plots in @fig-Fig1:
```{r plot-figure1, out.width=600, out.height=600}
#| label: fig-Fig1
#| fig-cap: "The number of residential parcels per DA in 2020 (top) and the rate of population per dwelling per DA (middle) and prevalence of Low-income cut-offs after-tax (LICO-AT) per DA (bottom) retrieved from the 2021 Canadian Census. All scales in quartiles. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/data/pop_descriptive_plots.tiff"))
```

@@

```{r plot-figure2, out.width=600, out.height=600}
#| label: fig-Fig2
#| fig-cap: "The locations of care destinations in Hamilton separated by the author-generated categories of: dependent-, errand-, grocery- and health- centric care categories. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/data/care_categories_plot.tiff"))
```

@@

```{r plot-figure3, out.width=600, out.height=600}
#| label: fig-Fig3
#| fig-cap: "The cumulative opportunity measure. The number of care destinations that can be reached, per DA, within 15 min walk. Visualised as 'caring' (based on the opportunity reachable number; Caring is approximately the median) and if complete (if all care categories can be reached; this is an arbitrary selection). Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_copp_weighted_plot.tiff"))
```

```{r plot-figure4, out.width=600, out.height=600}
#| label: fig-Fig4
#| fig-cap: "The diversity measure based on the _specific type_ of destination cumulative opportunity DA measure. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_entropy_copp_weighted_plot.tiff"))
```

```{r plot-figure5, out.width=600, out.height=600}
#| label: fig-Fig5
#| fig-cap: "A bivariate map of the cumulative opportuninities map (aggregated per category at DA - fig3) and the diversity measure (fig 4). Amenity diverse = diviersity index above the region median. Caring/Complete = accessibility value above the region median. The Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_bivariate_entropy_copp_plot.tiff"))
```

BONUS! Per capita spatial availability. This is not discussed in the methods. The spatial availability calculation assumes population is 'population density' and weighted opportunity count. With this map we can start thinking about per capita availability of opportunities. 
```{r plot-figure6, out.width=600, out.height=600}
#| label: fig-Fig6
#| fig-cap: "Per capita spatial availability of care opportunities. Uses weighted care opportunities, population density as competition, and 15 minute binary cut-off. Aggregates parcel level spatial availability at the DA level (by summation) and divides by the DA population. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_vvi_plot.tiff"))
```

@@

```{r plot-figure7, out.width=600, out.height=600}
#| label: fig-Fig7
#| fig-cap: "UNWEIGHTED -- The cumulative opportunity measure. The number of care destinations that can be reached, per DA, within 15 min walk. Visualised as 'caring' (based on the opportunity reachable number; Caring is approximately the median) and if complete (if all care categories can be reached; this is an arbitrary selection). Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_copp_plot.tiff"))
```

```{r plot-figure8, out.width=600, out.height=600}
#| label: fig-Fig8
#| fig-cap: "UNWEIGHTED -- The diversity measure based on the _specific type_ of destination cumulative opportunity DA measure. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_entropy_copp_plot.tiff"))
```
